Response to Reviewers

### 2021-11-15 Scripts/responses_rev_gcb.Rmd
misc

---
title: "Response to Reviewers"
output: pdf_document
---

# Edditor

@@

- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT. LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL**

Below you will find the response to both Reviewers 1 and 2. In addition, we would like to offer here an explanation of how our paper differs from recent published papers, including those authored by some of us.

@@

To identify the timeline we adopt the concept of *Time of Emergence* commonly used in oceanic physical and chemical studies (See Rodgers *et al.* 2015, Keller *et al.* 2014 and has just recently been used in ecological studies by Cheung *et al.* 2020 looking at ecological responses to marine heatwaves. As far as we are concerned, no other study has adopted this methodology to looking at the tyming of shifts in transbounary fisheries. On the other hand, to identify the intensity of the shift, we (partially) used methods developed by Palacios *et al.* and Sumaila *et al* (2020), published in *Ecology and society*. These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US tranboundary to manage four transboundary stocks. Here, we take this idea of threat point and expand the analysis globally to estimate the intensity that the shifts in transboundary stocks will have.

In short, the current manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed. We hope to have clarified the novelty of our research and how it differs from previous studies. The new information that we present here can be critical for fisheries management to adapt to shifts in the distribution of transboundary stocks under climate change as highlighted by both reviewers.

@@

- **Response:** We are glad that the reviewer found our manuscript to be interesting. We have revised the manuscript paying special attention to clarifying the parts the reviewer flagged. We thank the reviewer for his/her comments as they have strengthen our paper. Please find below direct responses to the reviewer's comments.

@@

- **Response:** Thank you for this comment. We acknowledge that we did not defined clearly our concept of stock in the previous version of the manuscript. The definition of stock has now been included in the revised manuscript in both the introduction and the methods as follows;

@@

  - **Response:** We understand the reviewer's concern and hope that our response have clarified this issue. However, it is almost certain that our definition of stock is influencing the results of our paper, just by that nature of the analysis. While a more fine scale definition (e.g., biologically-defined populations of stock will probably arose other results, not limiting the species to, at least, an EEZ could result in an overestimation of our results (See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118). This is a main caveat of our study and have made an effort to be very clear and upfront about it in line XX of the revised manuscript:

@@

- **Response:** We agree with the reviewer that and un-weighted centroid can severely skew the actual distribution of the species and influence our results. We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range. Specifically, the geographic centroid of each stock is indeed weighted by the abundance of the stock. The DBEM uses the species probability of occurrence to estimate the biomass of the species at each 0.5 x 0.5 gridcell as indicated from lines XX to XX of the methods. The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance (see below). This is, we determined the centroid of the stock using only those regions where abundance was higher (i.e., using the probability of occurrence of the DBEM). Moreover, we estimated the sensitivity of our results to the (arbitrarily) $per^{th} = 95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements. It reads as follows in line XX:

@@

![Top 5](../Figures/top_toe_map.png)

@@

- **Response:** This is indeed the definition of our stock. We do not differentiate between transboundary and straddling stocks as per defined in Palacios-Abanes *et al.* (2020). Please refer to our responses to the main comment.

@@

  - - Line XX; *Using a less conservative threshold (i.e., one s.d. representing a probability of 67% that the stock has shifted) results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks*

@@

  - - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*

@@

![New Figure 5](../Figures/Fig5n.png)

@@

  - - *This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*

@@

- line XX: **Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX). In some cases, the SSR of one EEZ is projected increase in more than double relative to their neighbors (e.g., from 10% to 30% SSR). Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively**

@@

  - - **Most EEZs are expected to see a relative small number of stocks shifting their SSR (Fig. \@ref(fig:toeFigure4)A). On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century. However, this result masks large changes, specially on island territories. By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Isl. (61%), Brazil (75%) and Kerguelen Isl., and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks. On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.**

@@

Comments:

@@

- **Response:** We agree with the reviewer and have included the SDG in the paper.

@@

17.16.1
Number of countries reporting progress in multi-stakeholder development effectiveness monitoring frameworks that support the achievement of the sustainable development goals

@@

- **Response:** Thank you for this comment. We have now changed the term to "*Transboundary Shift Index (TSI)*"

@@

- **Response:** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval. In fact, the distribution of some of our stocks does not follow this patter. Nevertheless, we believe our method responds well to other patterns (as long as they are considered one single stock unit) since the centroid of the stock's distribution is based on the "high biomass-level" grid cells. We have included a clarification regarding the stocks distribution in the figure legend of the revised manuscript.

@@

- **Response:** Red circles represent the years the TI is larger than the historical variation. These were quite faded in the original diagram and have made them clearer in the revised figure.

@@

- **Response:** The distributional centroid represent's the stocks centroid while "geographical" centroid is the EEZs centroid. We have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".

@@

@@

@@

@@

>>>>

---
title: "Response to Reviewers"
output:
  word_document: default
  pdf_document: default
---

# Editor

@@

- **Response: PLEASE PAY SPECIAL ATTENTION TO THIS POINT. LOOKS LIKE WE REALLY NEED TO CONVINCE THE EDITOR OUR RESEARCH IS NOVEL** Below you will find the response to both Reviewers 1 and 2. In addition, we would like to offer here an explanation of how our paper differs from recent published papers, including those authored by some of us.

@@

  To identify the timeline we adopt the concept of *Time of Emergence* commonly used in oceanic physical and chemical studies (See Rodgers *et al.* 2015, Keller *et al.* 2014 and has just recently been used in ecological studies by Cheung *et al.* 2020 looking at ecological responses to marine heatwaves. As far as we are concerned, no other study has adopted this methodology to look at the timing of shifts in transboundary fisheries. On the other hand, to identify the intensity of the shift, we (partially) used methods developed by Palacios *et al.* and Sumaila *et al* (2020), published in *Ecology and society*. These papers adopt the concept of threat point to evaluate the resilience of specific treaties celebrated by Canada-US transboundary to manage four transboundary stocks. Here, we take this idea of threat point and expand the analysis globally to estimate the intensity that the shifts in transboundary stocks will have.

  In short, our manuscript builds upon Palacios-Abrantes *et al.* 2020 *Scientific Reports* combining methods developed in Sumaila *et al.* 2020 and Palacios-Abrantes *et al.* 2020 *Ecology and Society* to address a question that neither Pinsky *et al.* 2018 *Science* nor Oremus *et al.* (2020) *Nature Sust.* addressed. We hope to have clarified the novelty of our research and how it differs from previous studies. The new information that we present here can be critical for fisheries management to adapt to shifts in the distribution of transboundary stocks under climate change as highlighted by both reviewers.

@@

- **Response:** We are glad that the reviewer found our manuscript to be interesting. We have revised the manuscript paying special attention to clarifying the parts the reviewer flagged. We thank the reviewer for his/her comments as they have strengthened our paper. Please find below direct responses to the reviewer's comments.

@@

- **Response:** Thank you for this comment. We acknowledge that we did not clearly define our concept of stock in the previous version of the manuscript. The definition of stock has now been included in the revised manuscript in both the introduction and the methods as follows;

@@

  - **Response:** We understand the reviewer's concern and hope that our responses have clarified this issue. However, it is almost certain that our definition of stock is influencing the results of our paper, just by that nature of the analysis. While a more fine scale definition (e.g., biologically-defined populations of stock will probably arose other results, not limiting the species to, at least, an EEZ could result in an overestimation of our results (See Ovando *et al.* https://www.pnas.org/content/118/23/e2025958118). This is a main caveat of our study and have made an effort to be very clear and upfront about it in line XX of the revised manuscript:

@@

- **Response:** We agree with the reviewer that an un-weighted centroid can severely skew the actual distribution of the species and influence our results. We have revised the manuscript to clarify that we did account for the heterogeneous distribution of fish stocks across their entire range. Specifically, the geographic centroid of each stock is indeed weighted by the abundance of the stock. The DBEM uses the species probability of occurrence to estimate the biomass of the species at each 0.5 x 0.5 gridcell as indicated from lines XX to XX of the methods. The centroid of a transboundary stock was determined by the average latitude and longitude ($lon_{ts}$) of the grid cells holding the specified percentile ($per^{th}$) of the projected transboundary stock abundance (see below). This is, we determined the centroid of the stock using only those regions where abundance was higher (i.e., using the probability of occurrence of the DBEM). Moreover, we estimated the sensitivity of our results to the (arbitrarily) $per^{th} = 95^{th}$ percentile threshold as described in lines XX and shown in Figure S6 of the supplements. It reads as follows in line XX:

@@

![Fig 4. Average year of range shift of the top five most valuable transboundary stocks of each EEZ between 2005 and 2014. Warm colors are indicative of an early shift  while cool colors represent a late shift. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue.](../Figures/Fig4.png)

@@

- **Response:** This is indeed the definition of our stock. We do not differentiate between transboundary, straddling nor highly-migratory stocks as per defined in Palacios-Abanes *et al.* (2020). Please refer to our responses to the main comment. Moreover, all of the catch and value data used represent fishing activities within countries’ EEZs.

@@

  - Line XX; *Using a less conservative threshold (i.e., one s.d. representing a probability of 67% that the stock has shifted) results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks*

@@

  - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*

@@

![New Figure 5](../Figures/Fig5.png)

@@

  - *This way, if, Canada and the US share a specific stock the SSR for Canada could be 20% and that of the US 80%, meaning that Canada holds 20% of the distribution of the stock within its waters and the US the other 80%.*

@@

- line XX: *Shifts in the SSR of the top 5 most valuable transboundary stocks of each country are projected to be 61 $\pm$ 18% by 2030 and 72 $\pm$ 17% by 2050 in 83% of the world's EEZs (Figure SX). In some cases, the SSR of one EEZ is projected to increase more than double relative to their neighbors (e.g., from 10% to 30% SSR). Some of these EEZs include Guatemala gaining from Mexico in the Pacific, Cyprus from Greece and Turkey in the Mediterranean Sea, and Russia from Norway and Japan in the Barents Sea and Western Pacific, respectively*

@@

  - *Most EEZs are expected to see a relatively small number of stocks shifting their SSR (Fig. \@ref(fig:toeFigure4)A). On average, EEZs are expected to see changes in SSR on 16% $\pm$ 10 of transboundary stocks increasing to 20% $\pm$ 11 by mid of the $21^{st}$ century. However, this result masks large variation in changes across EEZs. By early century, the EEZs of French Guiana (54%), South Africa (59%), Falkland Island (61%), Brazil (75%) and Kerguelen Island, and Pitcairn (both 100%) are projected to see changes in SSR of over 50% in their transboundary stocks. On the other hand, some island territories such as Anguilla are projected to see changes in as low as 2% of their transboundary stocks.*

@@

## Comments:

@@

- **Response:** We agree with the reviewer and have included the SDG 17 in the instrucution and discussion of the paper.

@@

- **Response: (NOTE: Help from COLETE here)** We have included the following in the discussion as a response to this comment:

  -- Line xx **The SDG target 17.16 (Enhance the global partnership for sustainable development) calls for collaborations between stakeholders to support the achievement of the SDGs in all countries, in particular developing countries [@SDGREF]. Such collaborations will be critical if we are to sustainably manage transboundary fisheries under climate change [@Pinsky:2018cb; @Oremus:2020id].**

@@

- **Response:** Thank you for this comment. We have now changed the term to *Transboundary Shift Index (TSI)*

@@

- **Response: (NOTE: Is the reviewer referring to multiple stocks within an EEZ??)** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval. In fact, the distribution of some of our stocks does not follow this pattern. Nevertheless, we believe our method responds well to other patterns (as long as they are considered one single stock unit) since the centroid of the stock's distribution is based on the "high biomass-level" grid cells. We have included a clarification regarding the stocks distribution in the figure legend of the revised manuscript.

@@

- **Response:** Red circles represent the years the TI is larger than the historical variation. They are all located in the graph along the black line.

@@

- **Response:** The distributional centroid represents the stocks centroid while "geographical" centroid is the EEZs centroid. We notice that this could be confusing and have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".

@@

- **Response: (NOTE: I AM UNSURE HOW TO RESPOND TO THIS COMMENT, ARE WE ASKED TO LINK THE UNCERTAINTY TO THE DECISION MAKING PROCESS?)**: Indeed uncertainty will be an inherit aspect of climate change and decision making. We have included a couple of limes in the discussion making a more evident linkage between the historical variability and the decision making process:

  - Line XX, *Therefore, decision-makers are facing the challenge of taking action under an uncertain future. One strategy for that is adapting scenario planning into fisheries management [@NOAA_scen_plan_2020]. This will allow for managers to integrate the uncertainty into planning for managing fisheries. For example, if stocks are expected to shift their distribution by 2036 $\pm$ 28 years (Fig. 2), managers could prepare for a scenario where stocks are currently changing and another where they will shift in 20 years.*

@@

- **Response:** Please see response to the previous comment

@@

- **Response:** Please see response to the Scenario planning comment.

@@

- **Response: (NOTE: Don't understand, what vectors?)**
